---
title: "Block and village-level VL - descriptive"
output:
  html_document:
    df_print: paged
---

```{r setup}

set.seed(1000)

library(tidyverse)
library(sf)
library(spdep)
library(raster)
library(rgeos)
library(patchwork)

theme_set(theme_minimal())
pal <- viridis::plasma(4)

# Fix default ggsave background
ggsave <- function(..., bg = 'white') ggplot2::ggsave(..., bg = bg)

sf::sf_use_s2(TRUE)

# Local data folder
datadir <- "~/VL/Data"

figdir <- "figures/descriptive"
```

```{r read_data}

v_raw <- read.csv(here::here(datadir,"KAMIS","Raw","village","village level data-Bi+Jh.csv"), header=TRUE) %>% 
  filter(state == "BIHAR") %>% 
  dplyr::select(vil_code, district, block, hsc, village, population, X2018) %>% 
  mutate(dist_blk = paste(district, block, sep = ":"))

v_shp_vl <- readRDS(here::here("data","analysis","v_shp_vl.rds")) %>% 
  st_transform(4326) %>% st_make_valid()
v_cent_vl <- readRDS(here::here("data","analysis","v_cent_vl.rds")) %>% 
  st_transform(4326)
b_shp_vl <- readRDS(here::here("data","analysis","b_shp_vl.rds")) %>% 
  st_transform(4326) %>% st_make_valid()
boundary <- readRDS(here::here("data","analysis","boundary.rds")) %>% 
  st_transform(4326)

v_shp_vl %>% 
  st_drop_geometry() -> v_df

```

```{r totals}

# Raw data
paste("2018 case count (raw):",sum(v_raw$X2018))

v_raw %>% 
  filter(X2018 > 0) %>% 
  summarise(n_blk = n_distinct(dist_blk),
            n_vil = n_distinct(vil_code))

# Analysis data
paste("2018 case count (analysis):",sum(v_shp_vl$X2018))

v_shp_vl %>% 
  st_drop_geometry() %>% 
  filter(X2018 > 0) %>% 
  summarise(n_blk = n_distinct(BLK_C_2011),
            n_vil = n_distinct(OBJECTID))

```

## Populations and village areas

```{r pops}

summary(v_df$pop_care_estimate[v_df$pop_care_estimate != 0])
summary(v_df$pop_raster_agg)

v_df %>% 
  summarise(n_zero_pop = sum(pop_raster_agg == 0),
            p_zero_pop = mean(pop_raster_agg == 0),
            n_zero_pop_wcases = sum(pop_raster_agg[X2018 > 0] == 0))

v_shp_vl %>% 
  st_transform(7759) %>% 
  st_area() %>% 
  units::set_units(km^2) -> area_km2 

summary(area_km2)
mean(as.numeric(area_km2 < 1*units::as_units("km^2")))
```

## Block level incidence

Summarise and plot observed incidence in 2018 by block:

```{r byblk}

summary(b_shp_vl$X2018_inc_blk*1e4)
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
# 0.0000  0.0000  0.0880  0.3015  0.3205  7.0854

ggplot(b_shp_vl, aes(x = X2018_inc_blk*1e4)) +
  geom_histogram() +
  labs(x = "Incidence per 10,000", y = "Frequency") -> inc_hist
inc_hist
ggsave(here::here(figdir,"hist_blkinc_2018.png"), height = 4, width = 5)

ggplot(b_shp_vl, aes(geometry = geometry, fill = X2018_inc_blk*1e4)) +
  geom_sf(data = boundary, fill = "grey") +
  geom_sf(col = NA) +
  scale_fill_viridis_c(trans = "sqrt") +
  labs(fill = "Incidence\nper 10,000") + 
  scale_x_continuous(breaks = c(84,86,88)) + 
  scale_y_continuous(breaks = c(25,26,27)) -> inc_map
inc_map
ggsave(here::here(figdir,"map_blk_inc_2018.png"), height = 4, width = 6)

inc_hist + inc_map + plot_annotation(tag_levels = "a")
ggsave(here::here(figdir,"blk_inc_2018.png"), height = 4, width = 10)
ggsave(here::here("figures/manuscript/Fig2.png"), height = 4, width = 10)

```

## Village level incidence

```{r vill_inc}

summary(v_shp_vl$X2018_inc*1000)

v_df %>% 
  filter(X2018_inc*1000 > 300) %>% 
  dplyr::select(DIST_NAME:V_NAME,X2018, pop_care_estimate:X2018_inc)

```

Plot observed incidence by village:

```{r plot_vill_inc}

ggplot(v_shp_vl, aes(fill = X2018*1000/pop_raster_agg)) +
  geom_sf(col = NA) +
  scale_fill_viridis_c(trans = "sqrt", na.value = "grey", direction = 1) + 
  labs(fill = "Incidence\nper 1000") +
  scale_x_continuous(breaks = c(84,86,88)) + 
  scale_y_continuous(breaks = c(25,26,27)) 
ggsave(here::here(figdir,"vil_inc_2018.png"), height = 4, width = 5)

```

### Block vs village incidence rates

Compare observed village and block incidence. Some village populations estimated as 0, so first need to replace resulting NaNs with 0s in village incidence. Then calculate the overall correlation and plot the scatter.

```{r blk_vil_inc}

v_shp_vl %>% 
  mutate(X2018_inc = replace_na(X2018/pop_raster_agg, 0)) -> v_shp_vl

cor(v_shp_vl$X2018_inc, v_shp_vl$X2018_inc_blk, method = "spearman") # 0.25

# Scatter plot (log scale; excluding 0 incidence)
v_shp_vl %>% 
  ggplot(aes(X2018_inc_blk, X2018_inc)) +
  geom_jitter(alpha = 0.3) +
  geom_abline() +
  labs(x = "Observed block incidence rate",
       y = "Observed village incidence rate") + 
  scale_x_continuous(trans = "log10") + 
  scale_y_continuous(trans = "log10")
ggsave(here::here(figdir,"obsvexp_vill_inc.png"), height = 4, width = 5)

v_shp_vl %>% 
  st_drop_geometry() %>% 
  mutate(X2018_exp = round(X2018_inc_blk*pop_raster_agg)) %>% 
  dplyr::select(OBJECTID, X2018_exp, X2018) %>% 
  pivot_longer(-OBJECTID) %>% 
  mutate(name = factor(name, 
                       levels = c("X2018_exp","X2018"), 
                       labels = c("Baseline prediction\n(Uniform incidence per block)","Observed"))) -> obs_exp_long
  
obs_exp_long %>% 
  ggplot(aes(as.factor(round(value)), fill = name)) +
  geom_bar(position = position_dodge(preserve = "single")) +
  scale_fill_manual(values = c(pal[1],"darkgrey")) +
  labs(fill = "", y = "No. villages", x = "Case count") +
  # ylim(0,1000) +
  scale_y_continuous(trans = "log2") +
  theme(legend.position = c(0.7,0.8)) -> bycat_all

bycat_all
ggsave(here::here(figdir,"obsvexp_bycats.png"), height = 4, width = 8)
ggsave(here::here("figures/manuscript/Fig3.png"), height = 4, width = 8)

```

## Spatial autocorrelation

Calculate Moran's I statistic for evidence of correlation between neighbouring blocks:

```{r block_autocorr}

# Specify cluster neighbour structure/weights
nb <- poly2nb(b_shp_vl)
lw <- nb2listw(nb, style = "B", zero.policy = TRUE)
  
bMC <- moran.mc(b_shp_vl$X2018_inc_blk, lw, nsim=999, zero.policy = TRUE)
bMC
plot(bMC, xlab = "Block-level incidence")

```

Also calculate across villages:

```{r village_autocorr}

# Specify cluster neighbour structure/weights
nb <- poly2nb(v_shp_vl)
lw <- nb2listw(nb, style = "B", zero.policy = TRUE)
  
vMC <- moran.mc(v_shp_vl$X2018_inc, lw, nsim=999, zero.policy = TRUE)
vMC
plot(vMC, xlab = "Village-level incidence")

```

```{r save_plot_I}

# Extract simulated Moran's I values
simI_blk <- bMC$res
simI_vil <- vMC$res

# Create a data frame for ggplot
moran_data <- data.frame(simI_blk = simI_blk,
                         simI_vil = simI_vil) 

moran_data %>%  
  ggplot(aes(x = simI_blk)) +
  geom_histogram() +
  geom_vline(aes(xintercept = bMC$statistic), linetype = "dashed") +
  labs(subtitle = "By block",
       x = "Simulated Moran's I",
       y = "Frequency") +
  theme_minimal() -> plot_bMC

moran_data %>%  
  ggplot(aes(x = simI_vil)) +
  geom_histogram() +
  geom_vline(aes(xintercept = vMC$statistic), linetype = "dashed") +
  labs(subtitle = "By village",
       x = "Simulated Moran's I",
       y = "Frequency") +
  theme_minimal() -> plot_vMC

plot_bMC + plot_vMC + plot_annotation(title = "Monte Carlo Simulation of Moran's I", tag_levels = "a")
ggsave(here::here("figures/manuscript/supplementary/FigS3_revised.png"), height = 4, width = 9, dpi = 300)

```

Calculate statistic across villages within each block, to compare strength of correlation between high/low incidence blocks:

```{r moran_byblk}

calc_moranMC_blk <- function(i){
  
  # Specify cluster neighbour structure/weights
  nb <- poly2nb(v_shp_vl[v_shp_vl$BLK_C_2011 == i,])

  lw <- nb2listw(nb, style = "B", zero.policy = TRUE)

  MC <- moran.mc(v_shp_vl$X2018_inc[v_shp_vl$BLK_C_2011 == i], lw, nsim=999, zero.policy = TRUE)
  return(c(I = MC$statistic, p = MC$p.value))

}

# Only run on non-zero blocks
blknms <- unique(v_shp_vl$BLK_C_2011[v_shp_vl$X2018_inc_blk > 0])
MC_byblk <- lapply(blknms, calc_moranMC_blk)

bind_rows(MC_byblk) %>% 
  mutate(blk_inc = b_shp_vl$X2018_inc_blk[b_shp_vl$X2018_inc_blk > 0]) %>%
  ggplot(aes(blk_inc*10000, I.statistic)) + 
  geom_hline(yintercept = 0, col = "grey") +
  geom_point(alpha = 0.5) +
  scale_x_continuous(trans = "sqrt") +
  labs(x = "Block incidence per 10,000", y = "Moran's I statistic from village incidence")
ggsave(here::here(figdir,"moran_I_byblk.png"), height = 5, width = 6)

bind_rows(MC_byblk) %>% 
  mutate(blk_inc = b_shp_vl$X2018_inc_blk[b_shp_vl$X2018_inc_blk > 0]) %>%
  ggplot(aes(blk_inc*10000, p)) + 
  geom_hline(yintercept = 0.01, col = "grey", lty = "dashed") +
  geom_point(alpha = 0.5) +
  scale_x_continuous(trans = "sqrt") +
  labs(x = "Block incidence per 10,000", y = "P-value of Moran's I from village incidence")
ggsave(here::here(figdir,"moran_p_byblk.png"), height = 5, width = 6)

```

## Covariates

Summarise village covariate values by VL affected:

```{r cov_table}

# Non-standardised covariate raster stack
covariate_stack <- readRDS(here::here("data","covariates","analysis","cov_stack_1km.rds"))

# Keep city travel time instead of distance to major road
covariate_stack <- raster::dropLayer(covariate_stack, "dist_majroad")
covariate_stack <- raster::dropLayer(covariate_stack, "built_ext")

# Drop population raster
names(covariate_stack)
stk <- covariate_stack[[-1]]

covs <- exactextractr::exact_extract(stk, v_shp_vl, fun = "mean")
names(covs) <- gsub("mean.","", names(covs))

v_df <- v_df %>% 
  dplyr::select(-elev:-ndvi_sd) %>% 
  bind_cols(covs) 

paste_range <- function(var, digits = 2){
  med <- round(median(var),digits)
  range <- paste(round(quantile(var, c(0.25,0.75)),digits), collapse = ", ")
  v <- paste0(med, " [", range,"]")
  return(v)
}

v_df %>% 
  mutate(Overall = TRUE,
         Affected = (X2018 > 0),
         Unaffected = (X2018 == 0)) %>% 
  pivot_longer(c("Overall","Affected","Unaffected")) %>% 
  filter(value) %>% 
  mutate(name = factor(name, levels = c("Overall","Affected","Unaffected"))) %>% 
  group_by(name) %>% 
  summarise(N = n(),
            pop = paste_range(pop_raster_agg, digits = 1),
            across(c(elev:ndvi_sd), paste_range)) %>% 
  column_to_rownames("name") %>%
  ungroup() -> tab1

tab1

write.csv(t(tab1), here::here("output","tab1.csv"), row.names = T)

```

```{r plot_covs_raster}

plot_cov <- function(i, nm){
  
  r <- subset(covariate_stack, i)
  r <- raster::mask(r, as_Spatial(boundary))
  
  r_spdf <- as(r, "SpatialPixelsDataFrame")
  r_df <- as.data.frame(r_spdf) %>% 
    rename(layer = 1)
  
  ggplot() +
    geom_raster(data = r_df, aes(x = x, y = y, fill = layer)) +
    scale_fill_viridis_c(trans = "identity", na.value = "white") + 
    labs(title = nm, fill = "", x = "", y = "") +
    theme(axis.text = element_blank()) +
    coord_quickmap() -> p
  
 return(p)
  
}

names <- c("Population","Elevation","Distance to water","Travel time to city",
                            "LST (mean)","LST (SD)", "NDVI (mean)", "NDVI (SD)")

covplots <- purrr::map2(1:nlayers(covariate_stack), names, plot_cov)

plotgrid <- gridExtra::grid.arrange(grobs = as.list(covplots))

ggsave(here::here("figures","covariates","covariates_all.png"), plotgrid, height = 10, width = 15, dpi = 250)

for (p in seq_along(covplots)){
  png(here::here("figures","covariates",paste0(names(covariate_stack)[p],".png")), height = 800, width = 1000, res = 200)
  print(covplots[p])
  dev.off()
}

```


```{r plot_covs_village}

v_df %>% 
  pivot_longer(c(pop_raster_agg,elev:ndvi_sd)) %>% 
  mutate(inc = replace_na(X2018_inc*1000, 0)) %>% 
  ggplot(aes(value, inc)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  facet_wrap(~name, scales = "free") + 
  labs(x = "Village-averaged covariate",
       y = "Village incidence per 1,000") -> cont_orig
cont_orig

v_df %>% 
  pivot_longer(c(pop_raster_agg,elev:ndvi_sd)) %>% 
  mutate(inc = replace_na(X2018_inc*1000, 0)) %>% 
  ggplot(aes(value, inc)) +
  geom_point(alpha = 0.3) +
  geom_smooth() +
  scale_x_continuous(trans = "log") +
  scale_y_continuous(trans = "log") +
  facet_wrap(~name, scales = "free") + 
  labs(x = "Village-averaged covariate",
       y = "Village incidence per 1,000") -> cont_trans
cont_trans

v_df %>% 
  pivot_longer(c(pop_raster_agg,elev:ndvi_sd)) %>% 
  mutate(inc = replace_na(X2018_inc > 0, FALSE)) %>% 
  ggplot(aes(value, inc)) +
  geom_boxplot(alpha = 0.3) +
  facet_wrap(~name, scales = "free") + 
  labs(x = "Village-averaged covariate",
       y = "Village incidence > 0") -> discr_orig
discr_orig

v_df %>% 
  pivot_longer(c(pop_raster_agg,elev:ndvi_sd)) %>% 
  mutate(inc = replace_na(X2018_inc > 0, FALSE)) %>% 
  ggplot(aes(value, inc)) +
  geom_boxplot(alpha = 0.3) +
  scale_x_continuous(trans = "log") +
  facet_wrap(~name, scales = "free") + 
  labs(x = "Village-averaged covariate",
       y = "Village incidence > 0") -> discr_trans
discr_trans
```